---
title: "chem_analysis_1"
output: html_document
---

```{r setup, include=FALSE}
```

## Load libraries
```{r chem}
library("tidyverse")
library("tidyr")
library("dplyr")
library("ggplot2")
library("readr")
library("RColorBrewer")
library("reshape2")
```

## Set up file pathways
```{r}
project.dir <- file.path("/Users/coin8046/Desktop/FiererLabNotebook_copy/mustard_microbiome/mustard_chem")
  rawdata.dir <- paste0(project.dir, "/00_raw_data")
  output.dir <- paste0(project.dir, "/01_outputs")
  fig.dir <- paste0(output.dir, "/figs")
  
if (!dir.exists(output.dir)) {dir.create(output.dir)}  
  if (!dir.exists(fig.dir)) {dir.create(fig.dir)}  
```

## Read in data
```{r}
chem.raw <- read.csv(paste0(rawdata.dir, "/GSL_data.csv"))

plant.raw1 <- read.csv(paste0(rawdata.dir, "/plant_data/Lab Notebook Mustard Project - growth stages_1.26.22.csv"))
plant.raw2 <- read.csv(paste0(rawdata.dir, "/plant_data/Lab Notebook Mustard Project - plant height.csv"))
plant.raw3 <- read.csv(paste0(rawdata.dir, "/plant_data/Lab Notebook Mustard Project - seed harvest.csv"))
plant.raw4 <- read.csv(paste0(rawdata.dir, "/plant_data/Lab Notebook Mustard Project - seed_weight.csv"))
```

## Format data
### chemistry data
```{r}
chem.gsl <- chem.raw %>% 
  select(-c("Location", "count", "Plate.Position", "X", "X.1")) %>% 
  mutate(treatment = gsub("\\-.*","", sample.name)) %>% 
  mutate(soil_type = parse_number(treatment)) %>% 
  mutate(sterility = gsub("^([[:alpha:]]*).*$","\\1",treatment))

chem.gsl[is.na(chem.gsl)] = 0
as.factor(chem.gsl$soil_type)


chem.gsl.long <- chem.gsl %>% 
  pivot_longer(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.", "X3MT..µmol.g.", "Indoles..µmol.g."), names_to = "gsl")
```

### plant data
```{r}

plant1.tmp <- plant.raw1 %>% 
  select(c("Pot.ID", "flower.buds.appear", "full.flower.appear", "seed.pods.appear", "no.flowers..pods.drying", "Spider.mites"))
  
plant3.tmp <- plant.raw3 %>% 
  select(c("Pot.ID", "total.seed.count", "Last.Seeds.Collected", "Root.Rhizo.Harvest")) %>% 
  mutate(total.seed.count = as.numeric(total.seed.count))

plant4.tmp <- plant.raw4 %>% 
  select(c("Sample.ID", "Total")) %>% 
  rename(total_seed_mass = Total) %>% 
  mutate(total_seed_mass = as.numeric(total_seed_mass))

plant.data.tmp <- plant1.tmp %>% 
  inner_join(plant3.tmp, by = "Pot.ID") %>% 
  inner_join(plant4.tmp, by = c("Pot.ID" ="Sample.ID"))

# cor.test(as.numeric(plant.data.tmp$total.seed.count), as.numeric(plant.data.tmp$total_seed_mass))

```

### plant height data
```{r}
## remove rows and columns with NAs
plant.raw2.1 <-  plant.raw2[!(plant.raw2$Pot.ID==""), ]
plant.raw2.2 <- plant.raw2.1[ , colSums(is.na(plant.raw2.1)) == 0]

## rename starting height column
colnames(plant.raw2.2)[2] <- "start"

plant_height <- plant.raw2.2 %>% 
  mutate_at(vars(-matches("Pot.ID")), as.numeric) %>% 
  mutate_at(vars(matches("height")),  ~ . - start) %>% 
  ## add metadata columns
  mutate(treatment = gsub("\\-.*","", Pot.ID)) %>% 
  mutate(soil_type = parse_number(treatment)) %>% 
  mutate(sterility = gsub("^([[:alpha:]]*).*$","\\1",treatment)) %>% 
  select(-start)
  
plant_height[is.na(plant_height)] = 0
as.factor(plant_height$soil_type)
plant_height.final <- plant_height[-c(31), ]

## filter to keep plants sent for chem sequencing
keep_ids.tmp <- c(chem.gsl$sample.name)
keep_ids <- gsub("\\..*","",keep_ids.tmp)

plant_height.filt <- plant_height.final %>% 
  filter(Pot.ID %in% keep_ids)

## tmp height data to merge with other plant data
plant_height.tmp <- plant_height.final %>% 
  select("Pot.ID", "height_9", "height_15")
 
### long formats
plant_height.long <- plant_height.final %>% 
  pivot_longer(cols = contains("height"), names_to = "height") %>% 
  mutate(timepoint = parse_number(height))

plant_height.long.filt <- plant_height.filt %>% 
  pivot_longer(cols = contains("height"), names_to = "height") %>% 
  mutate(timepoint = parse_number(height))
```

### reformat dates
```{r}
start_date <- as.Date("2021-03-31")
plant.data <- plant.data.tmp %>%
  mutate(flower.buds.appear = as.numeric(as.Date(flower.buds.appear, "%d%b%Y") - start_date)) %>% 
  mutate(full.flower.appear = as.numeric(as.Date(full.flower.appear, "%d%b%Y") - start_date)) %>% 
  mutate(seed.pods.appear = as.numeric(as.Date(seed.pods.appear, "%d%b%Y") - start_date)) %>%
  mutate(no.flowers..pods.drying = as.numeric(as.Date(no.flowers..pods.drying, "%d%b%Y") - start_date)) %>%
  mutate(Last.Seeds.Collected = as.numeric(as.Date(Last.Seeds.Collected, "%m/%d/%Y") - start_date)) %>%
  mutate(Root.Rhizo.Harvest = as.numeric(as.Date(Root.Rhizo.Harvest, "%m/%d/%Y") - start_date)) %>% 
  mutate(sample.name = paste0(Pot.ID, ".seed"))%>% 
  inner_join(plant_height.tmp, by = ("Pot.ID"))


plant_chem.data <- plant.data %>% 
  inner_join(chem.gsl, by = "sample.name") %>% 
  mutate(allyl_total = total_seed_mass*Allyl..µmol.g.)
 


 
```

### save data
```{r}
saveRDS(plant_chem.data, file = paste0(output.dir, "/plant.chem_data_11.10.RDS"))
write_csv(x = plant_chem.data, file = paste0(output.dir, "/plant.chem_data_23.5.3.csv"))

plant_chem.data <- readRDS(paste0(output.dir, "/plant.chem_data_11.10.RDS"))

plant_chem.long <- plant_chem.data %>% 
  pivot_longer(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.", "X3MT..µmol.g.", "Indoles..µmol.g."), names_to = "gsl")
```


## Plot data
### Glucosinolate data
#### gsl by treatment
```{r}

## plot all 4 gsl
p.treat <- ggplot(data = chem.gsl.long, mapping = aes(x = treatment, y = value))+
  geom_boxplot()+
  geom_point()+
  facet_wrap(~gsl, scales = "free_y")

p.treat

## plot one gsl
ggplot(data = chem.gsl, mapping = aes(x = treatment, y = `Allyl..µmol.g.`))+
  geom_boxplot()+
  geom_point()

```

```{r}
##### plotting settings
pair.pal2 <- c("#FFFF99", "#1F78B4", "#33A02C", "#E31A1C", "#FF7F00", "#6A3D9A", "#A6CEE3","#B2DF8A","#FB9A99","#FDBF6F","#CAB2D6")

pair.pal2nb <- c("#1F78B4", "#33A02C", "#E31A1C", "#FF7F00", "#6A3D9A", "#A6CEE3","#B2DF8A","#FB9A99","#FDBF6F","#CAB2D6")
pair.pal2nbG <- c("#1F78B4", "#33A02C", "#E31A1C", "#FF7F00", "#6A3D9A", "#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE")

palmuteGnb <- c("#332288", "#88CCEE", "#44AA99", "#117733", "#999933",  "#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE")
palmuteGnb2 <- c("#BEBEBE","#332288", "#88CCEE", "#44AA99", "#117733", "#999933",  "#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE")
palmuteG <- c("#BEBEBE", "#332288", "#88CCEE", "#44AA99", "#117733", "#999933",  "#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE","#BEBEBE") ##DDCC77",
palmute2 <- c("#BEBEBE", "#332288", "#88CCEE", "#44AA99", "#117733", "#999933", "#332288", "#88CCEE", "#44AA99", "#117733", "#999933", "#DDCC77")
```


#### gsl by soil type
```{r}
# New facet label names for supp variable
gsl.labs <- c("Allyl..µmol.g.","X3.Butenyl..µmol.g.","X3MT..µmol.g.","Indoles..µmol.g.")
names(gsl.labs) <- c("Allyl","Butenyl","3-methylthiopropyl","Indole")

chem.gsl.long.tmp <- chem.gsl.long
chem.gsl.long.tmp$gsl <- as.factor(chem.gsl.long.tmp$gsl)
levels(chem.gsl.long.tmp$gsl) <- c("Allyl","Butenyl","3-methylthiopropyl","Indole")

## plot all 4 gsl
# %>% subset(sterility != "S")
p.soil <- ggplot(data = chem.gsl.long.tmp, #%>% subset(sterility != "S"), 
                 mapping = aes(x = treatment, y = value, fill = treatment, group = treatment))+
  geom_boxplot()+
  #geom_violin(alpha = 0.5)+
  #geom_jitter(mapping = aes(x = treatment, y = value, color = treatment, group = treatment), size = 1.5)+
  #geom_point(mapping = aes(x = soil_type, y = value, color = treatment, group = treatment))+
  scale_color_manual(values=palmuteGnb, name = "Treatment")+
  scale_fill_manual(values=palmuteGnb, name = "Treatment")+
  facet_wrap(~gsl, scales = "free_y")+ #labeller = labeller(gsl= gsl.labs)
  xlab("Treatment")+
  ylab("Seed gsl concentration (µmol/g)")+
  scale_x_discrete(limits=c("M1","S1", "M2","S2", "M3","S3", "M4","S4", "M5", "S5"))+
  theme_bw(base_size = 20)
p.soil

ggsave(plot = p.soil, filename = paste0(fig.dir, "/gsl4xtreatnb_4.24.png"),width = 10, height = 6)

p2.soil <- ggplot(data = chem.gsl.long, mapping = aes(x = soil_type, y = value, group = soil_type))+
  geom_boxplot()+
  facet_wrap(~gsl, scales = "free_y")
p2.soil

### save
#ggsave(p.soil, filename = paste0(fig.dir, "/p.chem_soiltype.png"), width = 8, height = 6)

#ggsave(p2.soil, filename = paste0(fig.dir, "/p2.soiltype.png"), width = 8, height = 6)

## plot 1 gsl
p.allyl <- ggplot(data = chem.gsl %>% subset(sterility != "S"), mapping = aes(x = treatment, y = Allyl..µmol.g., fill = treatment, group = treatment))+
  geom_boxplot()+
  #geom_violin(alpha = 0.5)+
  #geom_jitter(mapping = aes(x = treatment, y = Allyl..µmol.g., color = treatment, group = treatment), size = 2)+
  scale_color_manual(values=palmuteGnb2, name = "Treatment")+
  scale_fill_manual(values=palmuteGnb2, name = "Treatment")+
  xlab("Soil inoculation treatment")+
  ylab("Allyl concentration (µmol/g)")+
  theme_bw(base_size = 20)
  #scale_x_discrete(limits=c("M1","S1", "M2","S2", "M3","S3", "M4","S4", "M5", "S5"))
p.allyl

ggsave(plot = p.allyl, filename = paste0(fig.dir, "/allylXtreatMAllylB_4.16.png"),width = 10, height = 6)
```



### total gsl by soil type
```{r}

## plot all total allyl
allylT.soil <- ggplot(data = plant_chem.data %>% subset(sterility != "B"), mapping = aes(x = soil_type, y = allyl_total, fill = treatment, group = treatment))+
  geom_boxplot()+
  scale_fill_manual(values=palmuteGnb)+
  xlab("Soil Type")+
  ylab("Total Allyl content in seeds (µmol)")+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))+
  theme_bw(base_size = 20)
allylT.soil

ggsave(plot = allylT.soil, filename = paste0(fig.dir, "/allylT.soil_11.10.png"),width = 10, height = 6)
```



#### gsl by sterility
```{r}
#color = as.factor(soil_type),
p.sterility <- ggplot(data = chem.gsl.long, mapping = aes(x = sterility, y = value,  group = sterility))+
  geom_boxplot()+
  facet_wrap(~gsl, scales = "free_y")

p.sterility

ggsave(p.sterility, filename = paste0(fig.dir, "/p.sterility.png"), width = 8, height = 6)

ggplot(data = chem.gsl, mapping = aes(x = sterility, y = `Allyl..µmol.g.`, color = as.factor(soil_type), group = sterility))+
  geom_boxplot()+
  geom_point()


```

### PCA of chemistry
### chem PCA / ordination
```{r}
### PCA
# https://www.datacamp.com/community/tutorials/pca-analysis-r

library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)

grp <- as.factor(chem.gsl$soil_type)

## all samples
chem.pca <- prcomp(chem.gsl[,3:6], center = TRUE,scale. = TRUE)
summary(chem.pca)
## plot PCA
p.chemPCA <- ggbiplot(chem.pca, ellipse = TRUE, groups = grp)+
  scale_colour_manual(values=pair.pal2)+
  theme_bw()
ggsave(plot = p.chemPCA, filename = paste0(fig.dir, "/chemPCAplot.png"))

## make df for PC1
chemPCs <- chem.pca$x %>% as.data.frame() %>% 
  rownames_to_column(var = "sampleID")
chemPC1 <- chemPCs %>% 
  dplyr::select(c("sampleID", "PC1"))

chem.gsl.pc <- chem.gsl %>% 
  cbind(chemPC1$PC1)

### plot PC1 values
p.PCA1 <- ggplot(data = chem.gsl.pc, mapping = aes(x = soil_type, y = chemPC1$PC1, color = treatment, group = treatment))+
  geom_boxplot()+
  scale_colour_manual(values=pair.pal2)+
  xlab("Soil Type")+
  ylab("PC1 (77%)")+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))
p.PCA1

ggsave(plot = p.PCA1, filename = paste0(fig.dir, "/PCA1xtreat.png"))
```



### plant growth and gsl
#### plant phenology
```{r}
#pair.pal <- brewer.pal(11, "Paired")
#display.brewer.pal(11, "Paired")
#pair.pal2 <- c("#FFFF99", "#1F78B4", "#33A02C", "#E31A1C", "#FF7F00", "#6A3D9A", "#A6CEE3","#B2DF8A","#FB9A99","#FDBF6F","#CAB2D6")   

plant_chem.long.tmp <- plant_chem.long
plant_chem.long.tmp$gsl <- as.factor(plant_chem.long.tmp$gsl)
levels(plant_chem.long.tmp$gsl) <- c("Allyl","X3 Butenyl","X3MT","Indoles")

## plot all 4 gsl
p.soil <- ggplot(data = chem.gsl.long.tmp %>% subset(sterility != "B"), mapping = aes(x = soil_type, y = value, fill = treatment, group = treatment))+
  geom_boxplot()+
  scale_fill_manual(values=palmuteGnb)+
  facet_wrap(~gsl, scales = "free_y")+ #labeller = labeller(gsl= gsl.labs)
  xlab("Soil Type")+
  ylab("Concentration (µmol/g)")+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))+
  theme_bw(base_size = 20)

###################################################
###################################################
## all gsl x flower appearing
p.flowerappear <- ggplot(data = plant_chem.long.tmp %>% subset(sterility != "B"), 
       mapping = aes(x = full.flower.appear, 
                     y = value))+
  geom_point(aes(color = treatment))+
  scale_colour_manual(values=palmuteGnb, name = "Treatment")+
  geom_line(stat="smooth", method = lm,
              size = 1.0, alpha = 0.5, color = "black")+
  facet_wrap(~gsl, scales = "free_y")+
    theme_bw(base_size = 20)+
  xlab("Flower appears (# days)")+
  ylab("Seed gsl concentration (µmol/g)")
  
p.flowerappear

ggsave(p.flowerappear, filename = paste0(fig.dir, "/p.flowerappear_12.13.png"), width = 8, height = 6)
###################################################
###################################################

## all gsl x last seed collected
p.lastseed <- ggplot(data = plant_chem.long, 
       mapping = aes(x = Last.Seeds.Collected, 
                     y = value))+
  geom_point(aes(color = treatment))+
  scale_colour_manual(values=pair.pal2)+
  geom_smooth(method='lm')+
  facet_wrap(~gsl, scales = "free_y")
p.lastseed

## all gsl x height
p.height1 <- ggplot(data = plant_chem.long, 
       mapping = aes(x = height_15, 
                     y = value))+
  geom_point(aes(color = treatment))+
  scale_colour_manual(values=pair.pal2)+
  #geom_smooth(method='lm')+
  facet_wrap(~gsl, scales = "free_y")
p.height1




ggsave(p.lastseed, filename = paste0(fig.dir, "/p.lastseed.png"), width = 8, height = 6)

## one gsl
ggplot(data = plant_chem.data, 
       mapping = aes(x = flower.buds.appear, 
                     y = `Indoles..µmol.g.`, 
                     color = treatment, 
                     group = as.factor(soil_type)))+
  geom_point()
  #geom_line()
```

#### seed production & gsl
```{r}
p.seedcount <- ggplot(data = plant_chem.long, 
       mapping = aes(x = total.seed.count, 
                     y = value))+
  geom_point(aes(color = treatment))+
  scale_colour_manual(values=pair.pal2)+
    geom_line(stat="smooth", method = lm,
              size = 1.0,
              alpha = 0.5)+
  facet_wrap(~gsl, scales = "free_y")
p.seedcount

###############3
###############
p.seedmass <- ggplot(data = plant_chem.long.tmp %>% subset(sterility != "B"), 
       mapping = aes(x = total_seed_mass, 
                     y = value))+
  geom_point(aes(color = treatment))+
  scale_colour_manual(values=palmuteGnb, name = "Treatment")+
  geom_line(stat="smooth", method = lm,
              size = 1.0, alpha = 0.5, color = "black")+
  facet_wrap(~gsl, scales = "free_y")+
  theme_bw(base_size = 20)+
  xlab("Total seed mass (g)")+
  ylab("Seed gsl concentration (µmol/g)")
p.seedmass

ggsave(p.seedmass, filename = paste0(fig.dir, "/p.seedmass_12.13.png"), width = 8, height = 6)
###############
###############

#plant_chem.data
p.allylT_seedmass <- ggplot(data = plant_chem.data, 
       mapping = aes(x = total_seed_mass, 
                     y = Allyl..µmol.g., 
                     color = treatment))+
  geom_point()+
  facet_wrap(~soil_type)+
  scale_colour_manual(values=pair.pal2)+
  geom_smooth(method = lm,se = FALSE)
  



ggsave(p.seedcount, filename = paste0(fig.dir, "/p.seedcount.png"), width = 8, height = 6)
ggsave(p.allylT_seedmass, filename = paste0(fig.dir, "/p.allylT_seedmass_11.10.png"), width = 8, height = 6)
 
```

#### spider mites and gsl
```{r}
p.spidermite <- ggplot(data = plant_chem.long %>% subset(Spider.mites != "I"), 
                       mapping = aes(x = Spider.mites, y = value,  group = Spider.mites))+
  geom_boxplot()+
  geom_jitter()+
  facet_wrap(~gsl, scales = "free_y")

p.spidermite
ggsave(p.spidermite, filename = paste0(fig.dir, "/p.spidermite.png"), width = 8, height = 6)

### t tests
mitesY <- plant_chem.data %>% subset(Spider.mites == "X")
mitesN <- plant_chem.data %>% subset(Spider.mites == "0") 

t.test(mitesY %>% subset(soil_type==1) %>% Allyl..µmol.g., mitesN$Allyl..µmol.g.)
t.test(mitesY$Indoles..µmol.g., mitesN$Indoles..µmol.g.)
t.test(mitesY$X3.Butenyl..µmol.g., mitesN$X3.Butenyl..µmol.g.)
t.test(mitesY$X3MT..µmol.g., mitesN$X3MT..µmol.g.)


plant.data %>% subset(Spider.mites == "X") %>% nrow()
plant.data %>% subset(Spider.mites == "0") %>% nrow()

ggplot(data = plant.data %>% subset(Spider.mites != "I"), 
                       mapping = aes(x = Spider.mites, y = total_seed_mass,  group = Spider.mites))+
  geom_point()+
  geom_boxplot()

ggplot(data = plant.data %>% subset(Spider.mites != "I"), 
                       mapping = aes(x = Spider.mites, y = total.seed.count,  group = Spider.mites))+
  geom_point()+
  geom_boxplot()

ggplot(data = plant.data %>% subset(Spider.mites != "I"), 
                       mapping = aes(x = Spider.mites, y = allyl,  group = Spider.mites))+
  geom_point()+
  geom_boxplot()
```

### seed production & treatment
```{r}
#x = soil_type, y = value, color = sterility, group = treatment
######################################################
######################################################
p.soil_seedmass <- ggplot(data = plant_chem.long %>% subset(sterility != "B"), 
                 mapping = aes(x = soil_type, 
                               y = total_seed_mass, 
                               #color = treatment, 
                               fill = treatment,
                               group = treatment))+
  geom_boxplot()+
  #scale_color_manual(values = palmuteGnb, name = "Treatment")+
  scale_fill_manual(values = palmuteGnb, name = "Treatment")+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))+
  xlab("Soil type")+
  ylab("Total seed mass (g)")+
  theme_bw(base_size = 30)
p.soil_seedmass

ggsave(p.soil_seedmass, filename = paste0(fig.dir, "/p.soil_seedmass_12.13.png"), width = 9, height = 6)



p.soil_flwr <- ggplot(data = plant_chem.long %>% subset(sterility != "B"), 
                 mapping = aes(x = soil_type, 
                               y = full.flower.appear, 
                               #color = treatment, 
                               fill = treatment,
                               group = treatment))+
  geom_boxplot()+
  #scale_color_manual(values = palmuteGnb, name = "Treatment")+
  scale_fill_manual(values = palmuteGnb, name = "Treatment")+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))+
  xlab("Soil type")+
  ylab("Flower appears (# days)")+
  theme_bw(base_size = 30)
p.soil_flwr

ggsave(p.soil_flwr, filename = paste0(fig.dir, "/p.soil_flwr_12.13.png"), width = 9, height = 6)

######################################################
######################################################


p.soil_seedcount <- ggplot(data = plant_chem.long, 
                 mapping = aes(x = soil_type, 
                               y = total.seed.count, 
                               color = treatment, 
                               group = treatment))+
  geom_boxplot()+
  scale_color_manual(values = pair.pal2)+
  scale_x_discrete(limits=c("1", "2", "3", "4", "5"))+
  xlab("Soil Type")+
  ylab("Total seed count per plant")+
  theme_grey(base_size = 18)
p.soil_seedcount


ggsave(p.soil_seedcount, filename = paste0(fig.dir, "/p.soil_seedcount.png"), width = 9, height = 6)
```

### plant height
```{r}
p.height_time <- ggplot(plant_height.long.filt, mapping = aes(x = timepoint, y = value, 
                                        color = treatment, group = Pot.ID)) +
  geom_line()+
  scale_colour_manual(values=pair.pal2)+
  facet_wrap(~treatment)
p.height_time 

ggsave(p.height_time, filename = paste0(fig.dir, "/p.height_time.png"), width = 8, height = 6)
```


### plant productivity / health
#### 
```{r}
## how do dif phenology dates correlate?
ggplot(plant.data, aes(x = seed.pods.appear, y = no.flowers..pods.drying))+
  geom_point()
cor.test(plant.data$flower.buds.appear, plant.data$no.flowers..pods.drying)

## phenology and seed mass?
ggplot(plant.data, aes(x = seed.pods.appear, y = total_seed_mass))+
  geom_point()
cor.test(plant.data$seed.pods.appear, plant.data$total_seed_mass)

## spider mite and...???


```

#### correlation matrix
```{r}
## http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
plant.data.df <- plant.data %>% 
  select(-c("Pot.ID", "sample.name"))

plant.data.df[is.na(plant.data.df)] <- 0

cormat <- round(cor(plant.data.df),2)

melted_cormat <- melt(cormat)
head(melted_cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

# Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(cormat)  
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# heatmap color palette hex codes
brewer.pal(n = 3, name = "RdBu")

# make heatmap pretty
p.plantcor <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "#67A9CF", high ="#EF8A62", mid = "#F7F7F7", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  #theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1), 
    axis.text.y = element_text(size = 12))
 #coord_fixed()

## plot with correlations in box
p.plantcor.lab <- p.plantcor +
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

ggsave(p.plantcor.lab, filename = paste0(fig.dir, "/p.plantcor.lab.png"), width = 8, height = 6)

ggsave(p.plantcor, filename = paste0(fig.dir, "/p.plantcor.png"), width = 8, height = 6)
```

## Stats
### should test for normality
```{r}

```

### ANOVAs
```{r}
## only gsl significant
x3mt.aov <- aov(`X3MT..µmol.g.` ~ treatment, data = chem.gsl)
summary(x3mt.aov)

aov.1 <- aov(total_seed_mass ~ treatment, data = plant_chem.data)
summary(aov.1)

library(rstatix)
library(car)
##not sig
allyl.aov <- aov(`Allyl..µmol.g.` ~ treatment, data = plant_chem.data) # %>% subset(treatment != "B") 
summary(allyl.aov)

Anova(allyl.aov, type="III")

allyl_acv <- anova_test(data = plant_chem.data, formula = Allyl..µmol.g. ~ treatment + total_seed_mass, type = 3, detailed = TRUE)
allyl_acv

```

### glm
```{r}
allyl.glm <- glm(`Allyl..µmol.g.` ~ treatment, data = plant_chem.data, 
                 ) # + total_seed_mass // sterility*soil_type + treatment*total_seed_mass
summary(allyl.glm)
```

### T tests
```{r}
## which soils have sig dif between microbial and sterile condition
t.test(data = chem.gsl %>% subset(soil_type == 2), 
       `X3MT..µmol.g.` ~ sterility)

wilcox.test(data = chem.gsl %>% subset(soil_type == 5), 
       `X3MT..µmol.g.` ~ sterility) 

## spider mites & gsl
t.test(data = plant_chem.data %>% subset(Spider.mites != "I"), 
       `Allyl..µmol.g.` ~ Spider.mites)
t.test(data = plant_chem.data %>% subset(Spider.mites != "I") %>% subset(soil_type == 5), 
       `Allyl..µmol.g.` ~ Spider.mites)

```

### Linear Models
```{r}
lm.1 <- lm(data = plant_chem.data,
   formula = Allyl..µmol.g. ~ Last.Seeds.Collected)
summary(lm.1)



```

### Correlations
```{r}
## allyl & seed mass
cor.test(plant_chem.data$Allyl..µmol.g., plant_chem.data$total_seed_mass)
# r = -0.5528698, p-value = 1.327e-08 

## allyl & flowering date
cor.test(plant_chem.data$Allyl..µmol.g., plant_chem.data$full.flower.appear)
# r = 0.2383573, p-value = 0.02289 

#### does seed mass correlate with total allyl
cor.test(plant_chem.data$allyl_total, plant_chem.data$total_seed_mass)
```

## average seed size
```{r}
plant_chem.dataP <- plant_chem.data %>% 
  mutate(avg_seed_size = total_seed_mass / total.seed.count)

cor.test(plant_chem.dataP$total_seed_mass, plant_chem.dataP$avg_seed_size)

cor.test(plant_chem.dataP$total_seed_mass, plant_chem.dataP$Allyl..µmol.g.)
```

