---
title: "mantel_tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up

### Load libraries
```{r}
library(mctoolsr)
library(plyr) 
library(tidyverse) 
library(vegan)
library(plotly)
library(readr)
library(viridis)

set.seed(10)
```


### Set up input and output directories
```{r Set up input and output directories, include = FALSE}
project.fp <- "/Users/coin8046/Desktop/FiererLabNotebook_copy/mustard_microbiome/mustard_16S/r_analyses"
clean_data.fp <- file.path(project.fp, "01_clean_data")
    figures_01_clean.fp <- file.path(clean_data.fp, "figures")
    outputs_01_clean.fp <- file.path(clean_data.fp, "outputs")

explr_02.fp <- file.path(project.fp, "02_exploration")
    figures_02.fp <- file.path(explr_02.fp, "figures")
    outputs_02.fp <- file.path(explr_02.fp, "outputs")
    
mantel_03.fp <- file.path(project.fp, "03_mantel")
    figures_03.fp <- file.path(mantel_03.fp, "figures")
    outputs_03.fp <- file.path(mantel_03.fp, "outputs")

# if (!dir.exists(mantel_03.fp)) {dir.create(mantel_03.fp, recursive = TRUE)}    
# if (!dir.exists(figures_03.fp)) {dir.create(figures_03.fp, recursive = TRUE)}
# if (!dir.exists(figures_03.fp)) {dir.create(figures_03.fp, recursive = TRUE)}
```

## Data set up 

### Read in data
```{r}
## data frames needed (generated in 02_exploration.c.Rmd)
# input_f2_samps5
# input_f2_samps6
# inputf2_rhizo.chem
# input_f2_rhizo.M
# input_f2_rhizo.S
# inputf2_root.chem
# input_f2_root.M
# input_f2_root.S
```


## Mantel tests

### Chemical mantel tests
Do differences in bacterial community composition correlate with differences in seed chemical composition?
#### Rhizosphere community :: seed chemistry
```{r}
## make chemical dataframe (with matching sample IDs)
chem.df1 <- inputf2_rhizo.chem$map_loaded %>% 
#   inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
#   column_to_rownames(var = "SampleID_2") %>% 
  dplyr::select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.", "X3MT..µmol.g.", "Indoles..µmol.g."))

## make total allyl dataframe (with matching sample IDs)
allyl.df1 <- inputf2_rhizo.chem$map_loaded %>% 
#   inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
#   column_to_rownames(var = "SampleID_2") %>% 
  dplyr::select(c("allyl_total"))

## separate out each individual gsl
chem.al <- chem.df1 %>% 
  dplyr::select("Allyl..µmol.g.")
chem.X3but <- chem.df1 %>% 
  dplyr::select("X3.Butenyl..µmol.g.")
chem.X3MT <- chem.df1 %>% 
  dplyr::select("X3MT..µmol.g.")
chem.ind <- chem.df1 %>% 
  dplyr::select("Indoles..µmol.g.")

## distance matrix for chemicals
dist.chem = dist(chem.df1)
dist.allyl = dist(allyl.df1)

dist.al <- dist(chem.al)
dist.X3but <- dist(chem.X3but)
dist.X3MT <- dist(chem.X3MT)
dist.ind <- dist(chem.ind)

## rhizosphere asv table (all rhizo samples)
rhizo_tab.r <- inputf2_rhizo.chem$data_loaded

## distance matrix for rhizosphere ASVs
rhizo_dm <- calc_dm(rhizo_tab.r, 'bray_sq_trans')
rhizo_dmJ <- calc_dm(rhizo_tab.r, 'jaccard')

## mantel test
rhizo_chem.mantel <- mantel(rhizo_dm, dist.chem)
rhizo_chem.mantel

rhizo_allyl.mantel <- mantel(rhizo_dm, dist.al)
rhizo_allylT.mantel <- mantel(rhizo_dm, dist.allyl)
```

#### Rhizosphere community :: seed chemistry (M & S treatments separate)
```{r}
### live microbial inoculation samples only

## chem dataframe
chem.df.Mrz <- input_f2_rhizo.M$map_loaded %>% 
  #inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  #column_to_rownames(var = "sample_id") %>% 
  dplyr::select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.","X3MT..µmol.g.", "Indoles..µmol.g."))

## all chem distance
dist.chem.Mrz = dist(chem.df.Mrz)

### PC1 distance
#dist.PC1.Mrz <- dist(chem.Mrz.all$PC1)

## rhizosphere asv table
rhizoM_tab.r <- input_f2_rhizo.M$data_loaded

## distance matrix for rhizosphere ASVs
rhizoM_dm <- calc_dm(rhizoM_tab.r, method = 'bray_sq_trans')
rhizoM_dmJ <- calc_dm(rhizoM_tab.r, method = 'jaccard')

## mantel test all chem
rhizo_M.mantel <- mantel(rhizoM_dm, dist.chem.Mrz, method = "spearman")
rhizo_M.mantel



### sterile inoculation samples only
chem.df.Srz <- input_f2_rhizo.S$map_loaded %>% 
  #inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  #column_to_rownames(var = "sample_id") %>% 
  dplyr::select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.","X3MT..µmol.g.", "Indoles..µmol.g."))

dist.chem.Srz = dist(chem.df.Srz)

## rhizosphere asv table
rhizoS_tab.r <- input_f2_rhizo.S$data_loaded

## distance matrix for rhizosphere ASVs
rhizoS_dm <- calc_dm(rhizoS_tab.r, method = 'bray_sq_trans')
rhizoS_dmJ <- calc_dm(rhizoS_tab.r, method = 'jaccard')

## mantel test
rhizo_S.mantel <- mantel(rhizoS_dm, dist.chem.Srz, method = "spearman")
rhizo_S.mantel
```

#### Spider mites mantel (plants with spider mites only)
```{r}
# data for plants with spider mites only
inputf2_rhizo.SMx <- filter_data(inputf2_rhizo.chem, filter_cat = "Spider.mites", keep_vals = "X")

# extract chem df
chem.SMrz <- inputf2_rhizo.SMx$map_loaded %>% 
  #inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  #column_to_rownames(var = "sample_id") %>% 
  dplyr::select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.","X3MT..µmol.g.", "Indoles..µmol.g."))

### all chem distance
dist.chem.SM = dist(chem.SMrz)

## extract rhizosphere asv table
SM.rhizo_tab <- inputf2_rhizo.SMx$data_loaded

## distance matrix for rhizosphere ASVs
SMrhizo_dm <- calc_dm(SM.rhizo_tab, method = 'bray_sq_trans')

mantel(SMrhizo_dm, dist.chem.SM)

```

#### Root community mantel tests
root community :: seed chemistry
```{r}
## make chemical dataframe (with matching sample IDs)
chem.df2 <- inputf2_root.chem$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  column_to_rownames(var = "sample_id") %>% 
  select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.", "X3MT..µmol.g.", "Indoles..µmol.g."))


## separate out each individual gsl
chem2.al <- chem.df2  %>% 
  select("Allyl..µmol.g.")
chem2.X3but <- chem.df2  %>% 
  select("X3.Butenyl..µmol.g.")
chem2.X3MT <- chem.df2  %>% 
  select("X3MT..µmol.g.")
chem2.ind <- chem.df2  %>% 
  select("Indoles..µmol.g.")

## distance matrix for chemicals
dist.chem2 = dist(chem.df2)

dist2.al <- dist(chem2.al)
dist2.X3but <- dist(chem2.X3but)
dist2.X3MT <- dist(chem2.X3MT)
dist2.ind <- dist(chem2.ind)

## root asv table
root_tab.r <- inputf2_root.chem$data_loaded

## distance matrix for root ASVs
root_dm <- calc_dm(root_tab.r)

## mantel tests
root_chem.mantel <- mantel(root_dm, dist.chem2)
root_chem.mantel

root_al.mantel <- mantel(root_dm, dist2.al)
root_al.mantel
root_X3but.mantel <- mantel(root_dm, dist2.X3but)
root_X3but.mantel
root_X3MT.mantel <- mantel(root_dm, dist2.X3MT)
root_X3MT.mantel
root_ind.mantel <- mantel(root_dm, dist2.ind)
root_ind.mantel
```

#### Root community :: seed chemistry (M & S treatments separate)
Are differences in root communities correlated with differences in seed chemistry? 
```{r}

### roots with live microbe treatment samples only
chem.df.Mrt <- input_f2_root.M$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  column_to_rownames(var = "sample_id") %>% 
  dplyr::select(c("Allyl..µmol.g..x", "X3.Butenyl..µmol.g..x","X3MT..µmol.g..x","Indoles..µmol.g..x"))

dist.chem.Mrt = dist(chem.df.Mrt)

## rhizosphere asv table
rootM_tab.r <- input_f2_root.M$data_loaded

## distance matrix for rhizosphere ASVs
rootM_dm <- calc_dm(rootM_tab.r, method = 'bray_sq_trans')

## mantel test
root_M.mantel <- mantel(rootM_dm, dist.chem.Mrt)
root_M.mantel


### sterile samples only
chem.df.Srt <- input_f2_root.S$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
  column_to_rownames(var = "sample_id") %>% 
  dplyr::select(c("Allyl..µmol.g..x", "X3.Butenyl..µmol.g..x","X3MT..µmol.g..x", "Indoles..µmol.g..x"))

dist.chem.Srt = dist(chem.df.Srt)

## rhizosphere asv table
rootS_tab.r <- input_f2_root.S$data_loaded

## distance matrix for rhizosphere ASVs
rootS_dm <- calc_dm(rootS_tab.r, method = 'bray_sq_trans')

## mantel test
root_S.mantel <- mantel(rootS_dm, dist.chem.Srt)
root_S.mantel
```

## PARTIAL MANTEL
#### add metadata and perform partial mantel
```{r}
# dfs to use (generated above)
## inputf2_rhizo.chem, input_f2_rhizo.M, input_f2_rhizo.S  
## rhizoM_dm, dist.chem.Mrz

# calc dist matrix for total_seed_mass; height 9 or 15
# dist() or vegdist()

# ditance matrix for rhizo community
rhizoM_dm <- calc_dm(input_f2_rhizo.M$data_loaded, method = 'bray_sq_trans')

# chem dataframe
chem.df1 <- input_f2_rhizo.M$map_loaded %>% 
#   inner_join(metadata, by = c("SampleID_3"= "Pot.ID")) %>% 
#   column_to_rownames(var = "SampleID_2") %>% 
  dplyr::select(c("Allyl..µmol.g.", "X3.Butenyl..µmol.g.", "X3MT..µmol.g.", "Indoles..µmol.g."))

## distance matrix for chemicals
dist.chem = dist(chem.df1)

transformed.chemdf1 <- sqrt(chem.df1)
dist.chemT = dist(transformed.chemdf1)

## distance matrices for metadata
dist.flwr <- dist(input_f2_rhizo.M$map_loaded$full.flower.appear)
dist.bolt <- dist(input_f2_rhizo.M$map_loaded$height_15)
dist.sd.mass <- dist(input_f2_rhizo.M$map_loaded$total_seed_mass)
# dist.flwrT <- dist(sqrt(input_f2_rhizo.M$map_loaded$full.flower.appear))
# dist.boltT <- dist(sqrt(input_f2_rhizo.M$map_loaded$height_9))


### mantel tests
mantel(rhizoM_dm, dist.chem)
mantel(rhizoM_dm, dist.bolt)
mantel.partial(rhizoM_dm, dist.chem, dist.bolt)
# mantel.partial(rhizoM_dm, dist.chem.Mrz, okabt_dist, method="pearson", permutations=999)
```


## M v S and plant productivity
Are differences in seed chemistry associated with differences in seed mass?
seed chemistry :: seed mass;
Are differences in seed mass correlated with differences in rhizosphere community?
rhizosphere community :: seed mass 
```{r}
## df of chemical metadata for rhizosphere samples
rhizo_meta <- inputf2_rhizo.chem$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID"))

### plot seed mass per condition
ggplot(data = rhizo_meta, aes(x = condition, y = total_seed_mass, 
                              color = condition))+
  geom_boxplot()+
  scale_colour_manual(values=pair.pal2)+
  scale_x_discrete(limits=c("B", "M1", "S1", "M2", "S2", "M3", "S3",
                            "M4", "S4", "M5", "S5"))

## chemistry dm
dist.chem

## seed mass dm
dist.seedmass <- dist(rhizo_meta$total_seed_mass)

## mantel test chem-seedmass
mantel.chem_seedmass <- mantel(dist.chem, dist.seedmass)
mantel.chem_seedmass


## mantel test seedmass - rhizosphere comm
mantel.seedmass_rhizo <- mantel(dist.seedmass, rhizo_dm)
mantel.seedmass_rhizo

```

#### Rhizo microbial and sterile samples separately
```{r}
### microbial samples only
M_rhizo_meta <- input_f2_rhizo.M$map_loaded #%>% 
  #inner_join(metadata, by = c("SampleID_3"= "Pot.ID"))
  
## distance matrix for seed mass
dist.seedmass <- dist(M_rhizo_meta$total_seed_mass)

## distance matrix for chem and rhizosphere ASVs
rhizoM_dm 
dist.chem.Mrz

## mantel test seed mass v chem
rhizo_M_meta.mantel <- mantel(dist.seedmass, dist.chem.Mrz)
rhizo_M_meta.mantel

## mantel test seed mass v rhizo
rhizo_M_sm.mantel <- mantel(rhizoM_dm, dist.seedmass)
rhizo_M_sm.mantel

## mantel test chem v rhizo
rhizo_M.mantel

##########################
##########################

### sterile samples only
S_rhizo_meta <- input_f2_rhizo.S$map_loaded #%>% 
  #inner_join(metadata, by = c("SampleID_3"= "Pot.ID"))

## distance matrix for seed mass
dist.seedmass.S <- dist(S_rhizo_meta$total_seed_mass)

## distance matrix for rchem and rhizosphere ASVs
dist.chem.Srz 
rhizoS_dm 

## mantel test seed mass v chem
rhizo_S_meta.mantel <- mantel(dist.seedmass.S, dist.chem.Srz)
rhizo_S_meta.mantel

## mantel test seed mass v rhizo
rhizo_S_sm.mantel <- mantel(rhizoS_dm, dist.seedmass.S)
rhizo_S_sm.mantel

## mantel test chem v rhizo
rhizo_S.mantel
```

#### Root microbial and sterile samples separately
```{r}
### microbial samples only
M_root_meta <- input_f2_root.M$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID"))
  
## distance matrix for seed mass
dist.seedmass.rt <- dist(M_root_meta$total_seed_mass)

## distance matrix for chem and rhizosphere ASVs
rootM_dm 
dist.chem.Mrt

## mantel test seed mass v chem
root_M_meta.mantel <- mantel(dist.seedmass.rt, dist.chem.Mrt)
root_M_meta.mantel

## mantel test seed mass v rhizo
root_M_sm.mantel <- mantel(dist.seedmass.rt, rootM_dm)
root_M_sm.mantel

## mantel test chem v rhizo
root_M.mantel

##########################
##########################

### sterile samples only
S_root_meta <- input_f2_root.S$map_loaded %>% 
  inner_join(metadata, by = c("SampleID_3"= "Pot.ID"))

## distance matrix for seed mass
dist.seedmass.Srt <- dist(S_root_meta$total_seed_mass)

## distance matrix for rchem and rhizosphere ASVs
dist.chem.Srt 
rootS_dm 

## mantel test seed mass v chem
root_S_meta.mantel <- mantel(dist.seedmass.Srt, dist.chem.Srt)
root_S_meta.mantel

## mantel test seed mass v rhizo
root_S_sm.mantel <- mantel(rootS_dm, dist.seedmass.Srt)
root_S_sm.mantel

## mantel test chem v rhizo
root_S.mantel
```

